\title{Fast Decompression Lucene Codec}
\author{Ivan Mamontov \\ 
Grid Dynamics}
\date{\today}


\documentclass[12pt]{article}

\usepackage{url} 

\begin{document}
\maketitle

\begin{abstract}

Sorted lists of integers are commonly used in Lucene's implementation of inverted index. Those lists are often compressed in-memory as a trade-off between memory footprint and access speed and CPU utilization. Thus, encoding and, more important, decoding of these lists consumes significant CPU time. 

We can use a SIMD instructions available in modern commodity processors to boost integer decompression performance. 
Key obstacle here is that Java/HotSpot doesn't provide access to SIMD instructions. Calling a JNI method from Java is rather expensive comparing to a simple C function call. Fortunately, for cases when native method just perform some encoding/decoding on a byte array, HotSpot has a private interface (it will likely to become a more standard extension) which adds a minimum overhead to the native routine.

In this talk we will show our prototype of Lucene Codec which uses a simple C library for compressing lists of integers using binary packing and SIMD instructions(\url{https://github.com/lemire/simdcomp}), which significantly improves decoding throughput.

\end{abstract}

%\section{Introduction}

%\section{Previous work}\label{previous work}

%\section{Results}\label{results}

%\section{Conclusions}\label{conclusions}

%http://www.stepanovpapers.com/CIKM_2011.pdf

\end{document}